---
title: "Lab 6 Comparing two means"
author: "Brett Vetter"
output: github_document
---

Researchers studying the number of electric fish species living in various parts of the Amazon basin were interested in whether the presence of tributaries affected the local number of electric fish species in the main rivers (Fernandes et al. 2004).

They counted the number of electric fish species above and below the entrance point of a major tributary at 12 different river locations.

The data is provided in your GitHub repository.

For each question below, write a sentence answering the question and show the code you used to come up with the answer, if applicable.

```{r}
library(tidyverse)

fish <- read_csv("chap12q19ElectricFish.csv")
fish_long <-
  pivot_longer(fish, speciesUpstream:speciesDownstream,
               names_to = "location",
               values_to = "species") %>%
  mutate(location = str_remove(location, c("species"))) %>%
  print()
```

## Question A

> What is the mean different in the number of species between areas upstream and downstream of a tributary? What is the 95% confidence interval of this mean difference. Show your code and write a sentence giving your answer

ANSWER

```{r}
t.test(formula = species ~ location, data = fish_long)
```

The mean difference in species between upstream and downstream areas is 1.84 species, and the 95% confidence interval for the means is (-4.58 8.25). 

## Question B

> Test the hypothesis that the tributaries have no effect on the number of species of electric fish.

ANSWER

From the T test we conducted we determined that we can reject the null hypothesis that the location has no effect on the number of species of electric fish. There was a significant difference in number of species between locations. 

## Question C

> State the assumptions that you had to make to complete parts (A) and (B). Create a graph to assess whether one of those assumptions was met.

ANSWER

```{r}
fish_long %>% 
  ggplot(aes(x = species)) +
  geom_histogram(
    aes(fill = location), 
    bins = 10, 
    alpha = 0.5, 
    position = "identity"
  ) 
```

My assumption for part A is that there IS a difference between the number of species in the two different locations (upstream, downstream). This would be the alternative hypothesis and it is the same as saying the difference in means is not equal to 0. My assumption for part B is that there is NO difference in the number of species in each location. This would be the Null hypothesis and it is the same as saying the difference in means is equal to 0. The histogram I created shows how the assumption from Part A was met.  