---
title: "Lab 6 Comparing two means"
author: "Brett Vetter"
output: github_document
---

Researchers studying the number of electric fish species living in various parts of the Amazon basin were interested in whether the presence of tributaries affected the local number of electric fish species in the main rivers (Fernandes et al. 2004).

They counted the number of electric fish species above and below the entrance point of a major tributary at 12 different river locations.

The data is provided in your GitHub repository.

For each question below, write a sentence answering the question and show the code you used to come up with the answer, if applicable.

```{r}
library(tidyverse)

fish <- read_csv("chap12q19ElectricFish.csv")
fish_long <-
  pivot_longer(fish, speciesUpstream:speciesDownstream,
               names_to = "location",
               values_to = "species") %>%
  mutate(location = str_remove(location, c("species"))) %>%
  print()

crab <- read_csv("chap15q27FiddlerCrabFans.csv")
```

## Question A

> What is the mean different in the number of species between areas upstream and downstream of a tributary? What is the 95% confidence interval of this mean difference. Show your code and write a sentence giving your answer

ANSWER

```{r}
t.test(formula = species ~ location, data = fish_long)
```

The mean difference in species between upstream and downstream areas is 1.84 species, and the 95% confidence interval for the means is (-4.58 8.25). 

## Question B

> Test the hypothesis that the tributaries have no effect on the number of species of electric fish.

ANSWER

From the T test we conducted we determined that we can reject the null hypothesis that the location has no effect on the number of species of electric fish. There was a significant difference in number of species between locations. 

## Question C

> State the assumptions that you had to make to complete parts (A) and (B). Create a graph to assess whether one of those assumptions was met.

ANSWER

```{r}
fish_long %>% 
  ggplot(aes(x = species)) +
  geom_histogram(
    aes(fill = location), 
    bins = 10, 
    alpha = 0.5, 
    position = "identity"
  ) 
```

My assumption for part A is that there IS a difference between the number of species in the two different locations (upstream, downstream). This would be the alternative hypothesis and it is the same as saying the difference in means is not equal to 0. My assumption for part B is that there is NO difference in the number of species in each location. This would be the Null hypothesis and it is the same as saying the difference in means is equal to 0. The histogram I created shows how the assumption from Part A was met.  

## ANOVA

Fiddler crabs are so called because males have a greatly enlarged "major" claw, which is used to attract females and to defend a burrow.

Darnell and Munguia (2011) recently suggested that this appendage might also act as a heat sink, keeping males cooler while out of the burrow on hot days.

To test this, they placed four groups of crabs into separate plastic cups and supplied a source of radiant heat (60-watt light bulb) from above. The four groups were intact male crabs, male crabs with the major claw removed; male crabs with the other (minor) claw removed (control); and intact female fiddler crabs.

They measured the body temperature of crabs every 10 minutes for 1.5 hours. These measurements were used to calculate a rate of heat gain for every individual crab in degrees C/log minute. Rates of heat gain for all crabs are provided in the accompanying data file.

### Question D

Graph the distribution of body temperatures for each crab type:

```{r}
crab %>% 
  ggplot(aes(x = bodyTemperature )) +
  geom_histogram(
    aes(fill = crabType), 
    bins = 15, 
    alpha = 0.5, 
    position = "identity",
    na.rm = TRUE
  ) 
```

### Question E

Does body temperature varies among crab types? State the null and alternative hypothesis, conduct and ANOVA, and interpret the results.

```{r}
aov_crabs <-
  aov(bodyTemperature ~ crabType, data = crab)
aov_crabs 

summary(aov_crabs)
```
The The null hypothesis is that the body temp does not vary at all between the crab types. The alternative hypothesis is that at least one crab type varies in body temp compared to the other types. Because the p value is less than 0.05 we can reject the null, and because under the alternative hypothesis F would be greater than 1 and the F was much greater than 1, this tells us that body temp in at least one crab type does significantly vary. 